Causal Compatibility Inequalities Admitting of Quantum Violations in the Triangle Structure

### 2017-09-08 documents/TriangleScenarioArXiv/ts.tex
substantial edits to tone, a few paragraphs removed.

\begin{document}
    \title{Causal Compatibility Inequalities Admitting of Quantum Violations in the Triangle Structure}
    \author{Thomas C. Fraser}
    \email{tcfraser@tcfraser.com}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada, N2L 2Y5}
    \affiliation{University of Waterloo, Waterloo, Ontario, Canada, N2L 3G1}
    \author{Elie Wolfe}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada, N2L 2Y5}
    % \email{ewolfe@perimeterinstitute.ca}
    \date{\today}
    \begin{abstract}
        It has long been recognized that certain quantum correlations are incompatible with particular assumption about classical causal structure. Given a causal structure of unknown classicality, the presence of such correlations certifies the non-classical nature of the causal structure in a device independent fashion. In the context of two parties sharing a common resource, these non-classical correlations are also known as non-local correlations. Any constraint which is satisfied by all correlations which \textit{are} classically compatible with a given causal structure defines a causal compatibility criterion. Such criteria were recently were recently derived for the Triangle structure [\href[pdfnewwindow]{http://arxiv.org/abs/1609.00672}{arXiv:1609.00672}] in the form of polynomial inequalities, begging the question: do any of those inequalities admit violation by quantum correlations? Numerical investigation suggests not, and we further conjecture that the set of correlations admitted by the classical Triangle structure is equivalent to the set of correlations admitted by its quantum generalization whenever the three observable variables are binary. Our main contribution in this work, however, is the derivation of new causal compatibility inequalities for the Triangle structure which \textit{do} admit quantum violation. We conclude by considering the possibility of quantum resources potentially qualitatively different from those known previously.
    \end{abstract}
    \maketitle
    % \tableofcontents
    % \setlength{\parskip}{1em}

@@

    From a foundational prospective, the most robust demonstrations of quantum phenomena with no classical emulation have involved Bell inequalities~\cite{Bell_1964}. Originally, Bell inequalities were derived as a way to show that no hidden variable theory could ever account for quantum mechanics; in this sense Bell inequalities are a response to the famous EPR paradox~\cite{EPR_Orig}. The enumeration of Bell inequalities has since become a widespread systematic way of demonstration the \textit{non-classicality} of a given observation. More recently it has been appreciated that Bell inequalities can understood as consequences of \term{causal inference}~\cite{Wood_2012}. Causal inference is concerned with classifying observations into those which can and cannot be explained by a hypothesized causal structure. The abstract nature of causal inference is responsible for its presence in numerous scientific fields including machine learning and biology~\cite{Pearl_2009,Pearl_2009_tr}. Causal compatibility inequalities, such as Bell and Instrumental inequalities, characterize the space of observations that are compatible with a hypothesized causal structure, albeit the characterization offered by practically derivable inequalities is often only an approximation. To derive the traditional Bell inequalities from causal inference one starts with a (classical) causal structure known as the Bell structure, as depicted in \cref{fig:bell_structure}. The fundamental Bell structure involves non-communicating parties making measurements on some hidden shared resource $\la$, where the measurement outcomes ($A, B$) are presumed to be stochastic functions of the the local choices of measurement settings ($S_A, S_B$) and the shared resource $\la$. Quantum non-classicality in the Bell structure has been thoroughly studied since Bell's original work~\cite{Brunner_2013}. More complex structures, however, such as the correlation scenarios proposed by Fritz~\cite{Fritz_2012,Fritz_2014}, are much less understood. Here we investigate one particular causal structure: the Triangle structure (\cref{fig:triangle_structure}).

    The Triangle structure (\cref{fig:triangle_structure}) is a causal structure composed of $3$ parties labeled $A, B, C$ arranged in a triangular configuration while pair-wise sharing hidden/latent variables $X, Y, Z$. It has been extensively studied previously (see~\cite[Fig. 1]{Steudel_2010},~\cite[Fig. 6]{Chaves_2014},~\cite[Fig. 8]{Branciard_2012},~\cite[Fig. 8, App. E]{Henson_2014},~\cite[Fig. 3]{Fritz_2012},~\cite[Fig. 4]{Weilenmann_2016},~\cite[Fig. 1]{Inflation}, $\ldots$). An overview of some milestone results is provided in \cref{sec:triangle_structure}. Identifying causal compatibility inequalities for this configuration has been seen as particularly challenging~\cite{Branciard_2012}. Further identifying causal compatibility inequalities of such high resolution that such inequalities can be violated by quantum-accessible distributions has remained out-of-reach for the Triangle structure.

    This work is the first, therefore, to find causal compatibility inequalities for the Triangle structure that are able to be violated by quantum-accessible distributions. This accomplishment was made possible through the combination of two previous developments. The first, an insight made by \citet{Fritz_2012} regarding the ability to re-interpret the Bell structure as a portion of the Triangle structure. The second, a new framework for solving causal inference problems developed by Wolfe \emph{et al.}~\cite{Inflation} called \term{The Inflation Technique}. Ultimately, this work serves as a validation that the Inflation Technique is sensitive enough to offer new insights into quantum non-classicality. Moreover, these new inequalities offer an avenue for recognizing previously unknown forms of non-classicality. The authors attempts to find such novel resources were met with only partial success, suggesting clear refinements for future exploration.

    The manuscript is organized into the following sections: \Cref{sec:causal_compatibility} recalls important notions from causal inference theory and sets up the notation to be used. \Cref{sec:bell_structure} offers a summary of the popular Bell structure and associated inequalities. \Cref{sec:triangle_structure} discusses the Triangle structure and provides an overview of existing research; identifying its stark differences from the Bell structure and motivating why the Triangle structure is worth studying. \Cref{sec:fritz_distribution} defines and discusses a singularly-quantum correlation first conceived of by \citet{Fritz_2012}, which we term the \term{Fritz distribution}. The Fritz distribution was proven to be non-classical without the use of inequalities --- this report fills in the desired missing inequalities. A sample of these inequalities are presented in \cref{sec:found_inequalities}, specifically \cref{eq:ww_ineq,eq:web_inequality,eq:symmetic_web_inequality}. Finally, \cref{sec:violations_noise} introduces a method for parameterizing the space of quantum-accessible distributions which we used to numerically optimize against our derived inequalities in search of stronger types of non-classicality. Additionally, the Fritz distribution is subjected to noise in order to measure the robustness of the derived inequalities. \Cref{sec:conclusions} concludes with proposed avenues for further research.

    \Cref{sec:main_appendix} briefly summaries the Inflation Technique in the specific context of this work. Although the summary presented here is designed to be self-standing, a much more pedagogical introduction is offered by the original work~\cite{Inflation}. \Cref{sec:deriving_inequalities} demonstrates how the Inflation Technique was used to derive causal compatibility inequalities for the Triangle structure that are violated by quantum-accessible distributions.

@@

    Henceforth, we will utilize a number of familiar notions from graph theory and denote them accordingly. The \term{parents of a node} $n \in \nodes$ are all nodes which point directly into $n$, i.e. $\Pa[\graph]{n} \defined \bc{m \mid m \to n}$. Similarly defined are the \term{children of a node} $\Ch[\graph]{n} \defined \bc{m \mid n \to m}$. Recursively defined are the \term{ancestors of a node} $\An[\graph]{n} \defined \bigcup_{i\in\mathbb{N}} \Pa[\graph][i]{n}$ where $\Pa[\graph][i]{n} \defined \Pa[\graph]{\Pa[\graph][i-1]{n}}$ and $\Pa[\graph][0]{n} = n$ and the \term{descendants of a node} $\De[\graph]{n} \defined \bigcup_{i\in\mathbb{N}} \Ch[\graph][i]{n}$ where $\Ch[\graph][i]{n} \defined \Ch[\graph]{\Ch[\graph][i-1]{n}}$ and $\Ch[\graph][0]{n} = n$. Finally, we extend this notation to a subset of nodes $N \subseteq \nodes$ by performing a union over elements. As an example, the parents of the nodes $N \subseteq \nodes$ are denoted $\Pa[\graph]{N} = \bigcup_{n \in N} \Pa[\graph]{n}$.

@@

    Now that the notion of a causal structure $\graph$ is defined along with some standard graph-theoretic notation, it is now possible to precisely define the notion of compatibility between $\graph$ and a probability distribution $\prob[\nodes]$ defined over the nodes of $\graph$. Intuitively, the absence of certain causal connections within $\graph$ should limit the set of possible observations that can be made on $\graph$. Specifically, the causal structure $\graph$ hypothesizes that if each variable $n \in \nodes$ is causally influenced by its parents $\Pa[\graph]{n}$. Therefore, if each variable is conditioned on its parentage, the probability distribution $\prob[\nodes]$ should factorize accordingly:
    \[ \prob[\nodes] = \prod_{n \in \nodes} \prob[n \mid \Pa[\graph]{n}] = \prob[n_1 \mid \Pa[\graph]{n_1}] \times \cdots \times \prob[n_k \mid \Pa[\graph]{n_k}] \eq \label{eq:compatibility_as_factorization}\]
    % , except its own descendants when conditioned over $\Pa[\graph]{n}$. This is known as the \term{local Markov property}:\footnote{Here $A \indep B \mid C$ denotes statistical independence between $A$ and $B$ conditional on $C$ where $A,B,C \subseteq \nodes$, i.e. for a given probability distribution $\prob[\nodes]$, $\prob[AB \mid C] = \prob[A \mid C]\prob[B \mid C]$.}
    % \[ n \indep \br{\nodes \setminus \De[\graph]{n}} \mid \Pa[\graph]{n} \quad \forall n \in \nodes \eq \label{eq:local_markov_property} \]
    If a given distribution $\prob[\nodes]$ defined over \textit{all} of the nodes $\nodes$ of $\graph$ satisfies \cref{eq:compatibility_as_factorization}, then $\prob[\nodes]$ is said to be \term{compatible with $\graph$}. The conditional distributions in \cref{eq:compatibility_as_factorization} (i.e. $\{\prob[n \mid \Pa[\graph]{n}] \mid n \in \nodes\}$) are referred to as a set of \term{causal parameters for $\graph$}. If a distribution $\prob[\nodes]$ cannot be factorized according to \cref{eq:compatibility_as_factorization}, $\prob[\nodes]$ is said to be \term{incompatible} with $\graph$. The compatibility of a distribution $\prob[\nodes]$ with a causal structure $\graph$ pursuant to \Cref{eq:compatibility_as_factorization} is equivalent to the requirement that $\prob[\nodes]$ obeys all of the conditional independence constraints implied by the \textit{d-separation relations} between variables/nodes in $\graph$~\cite{Pearl_2009}. When one is specified with a joint distribution $\prob[\nodes]$ defined over all nodes of a causal structure $\graph$, it is possible to completely determine whether or not $\prob[\nodes]$ is compatible with $\graph$ by computing the causal parameters induced by $\prob[\nodes]$ and checking the equality of \cref{eq:compatibility_as_factorization}.

    A challenge is presented when one is supplied with a partial observation, i.e. a joint distribution $\prob[\nodes_{O}]$ where $\nodes_{O} \subset \nodes$ is some subset of variables referred to as the \term{observable nodes} $\nodes_O$. In such cases, $\prob[\nodes_{O}]$ does not induce a unique set of causal parameters for $\graph$ and \cref{eq:compatibility_as_factorization} can not be verified by direction calculation. Instead, compatibility between $\graph$ and $\prob[\nodes_{O}]$ depends on the existence or non-existence of a set of causal parameters for $\graph$ such that $\prob[\nodes_{O}] = \prod_{n \in \nodes_{O}} \prob[n \mid \Pa[\graph]{n}]$. The complementary, unobservable nodes are termed \term{latent nodes} $\nodes_L = \nodes \setminus \nodes_{O}$ representing hidden random variables that are hidden either by some fundamental process or cannot be measured due to other limitations.

    There are several approaches to tackling the compatibility problem when dealing with latent variables; there are two common approaches worth mentioning here. The first is to recognize that the conditional independence constraints implied by d-separation relations over the \textit{observable} variables form a family of necessary constraints for $\prob[\nodes_{O}]$. The failure to satisfy one of these observable independence relations immediately qualifies $\prob[\nodes_{O}]$ as incompatible with $\graph$. Often, the observable d-separation relations implied by a causal structure form an \textit{insufficient} set of constraints for compatible distributions. Such causal structures, which include the Bell structure, have been termed \textit{interesting}~\cite{Henson_2014}. For interesting causal structures, stronger techniques must be used~\cite{Inflation}. The primary causal structure of this paper is the Triangle structure --- one of the smallest interesting causal structures~\cite{Henson_2014}. An alternative approach requires the cardinality (number of outcomes) of the latent variables to be fixed. In most realistic scenarios, the nature of the latent variables is completely unspecified. More powerful approaches find constraints for $\prob[\nodes]$ in the form of probabilistic inequalities defined over $\prob[\nodes_{O}]$ and its marginals which constraint the set of compatible probability distributions. We term these constraints \term{causal compatibility inequalities}\footnote{We refer to these inequalities as ``causal compatibility inequalities'' instead of Bell inequalities for two reasons. Firstly, ``Bell inequalities'' usually are associated specifically with the Bell structure. Secondly, the inequalities derived in this work are fundamentally distinct from a typical Bell inequality in that these inequalities are \textit{polynomial} over $\prob[\nodes_{O}]$ instead of \textit{linear}.}. If a probability distribution $\prob[\nodes_{O}]$ happens to \textit{violate} any causal compatibility inequality, then that distribution deemed \textit{incompatible}. \Cref{sec:main_appendix} provides a brief discussion of existing techniques used to derive these inequalities.

    Unfortunately, a singular inequality cannot prove that a given distribution is \textit{compatible}, only \textit{incompatible}. Instead, a \textit{complete characterization} of compatibility would consist of a complete set of all valid causal compatibility inequalities; satisfaction of a complete set of inequalities does prove compatibility. Currently however, it is unknown how to obtain a complete characterization for any given causal structure, including the Triangle structure.

    From the perspective of identifying quantum non-classicality, a causal structure $\graph$ adopts the role of a classical hypothesis. Therefore, non-classicality becomes synonymous with incompatibility: if a distribution $\prob[\nodes_{O}]$ is incompatible, then it is non-classical. Henceforth, we will use these two terms interchangeably. From a resource standpoint, if the non-classicality of $\prob[\nodes_{O}]$ can be witnessed by an inequality $I$, but $\prob[\nodes_{O}]$ can be implemented using quantum states and measurements, then the causal compatibility inequality $I$ represents a physical task where quantum resources out-perform classical resources.

    \section{Bell Structure}
    \label{sec:bell_structure}
    This section aims to define the Bell structure and summarize the main results associated with its non-classical character. The purpose of this section is to equip an unfamiliar reader with the pertinent background and also to establish a context for which to draw comparisons between the advancements made toward understanding non-classicality of the Bell structure and the analogous results obtained for Triangle structure in the remainder of this work. Readers familiar with the Bell structure (sometimes called Bell scenario/experiment) will experience little obstacles proceeding to \Cref{sec:triangle_structure}.

    The Bell structure (\cref{fig:bell_structure}) refers to an iconic causal structure involving two distant observers which make local measurements $A, B$ determined by measurement settings $S_A, S_B$ on a shared latent resource $\la$~\cite{Brunner_2013}. The resulting observed correlations form a conditional probability distribution $\prob[AB|S_AS_B]$. Subject to the notions of compatibility presented in \cref{sec:causal_compatibility}, this distribution is compatible with \cref{fig:bell_structure} if and only if there exists some functional specification of $\la$ such that there exists a distribution $\prob[AB|S_AS_B,\la]$ such that $\prob[AB|S_AS_B] = \sum_{\la} \prob[AB|S_AS_B,\la]$ and also satisfies the following implementation of \cref{eq:compatibility_as_factorization}:
    \[ \prob[AB|S_AS_B,\la] = \prob[A|S_A,\la]\prob[B|S_B,\la] \]
    Any observed distribution $\prob[AB|S_AS_B]$ which fails to be explained by the classical causal hypothesis of the Bell structure is appropriately termed non-classical. Often, distributions incompatible with the Bell structure are referred to as non-local because the Bell structure indicates that the latent resource is shared locally and measurements are made locally. In particular,~\citet{Bell_1964} demonstrated that there exists distributions that are non-classical yet are attainable via quantum resources and measurements. If one does not wish to accept fine-tuning arguments, one must accept that quantum correlations can be non-local as no variation to the latent nodes can explain the same correlations~\cite{Wood_2012}.\footnote{This is unlike the Triangle structure wherein a global yet latent common cause is always an acceptable explanation for any correlation and does require fine-tuning. Therefore, we never refer to non-classical correlations in Triangle Structure as non-local.} A distribution is accessible with quantum resources and measurements if there exists a bipartite quantum state $\rho_{AB}$ of arbitrary dimension and measurement operators $M_{A|S_A}, M_{B|S_B}$ conditional indexed by the measurement settings such that $\prob[AB|S_AS_B] = \Tr\bs{\rho_{AB}M_{A|S_A}\otimes M_{B|S_B}}$. The non-classicality of quantum distributions for the Bell structure can be demonstrated through the use of Bell inequalities such as the CHSH inequality (\cref{eq:CHSH_ineq_orig}) which constrain the correlations between binary variables $A$ and $B$ for classical distributions~\cite{CHSH_Original}.
    \[ \ba{\p A\p B | S_{\p{A}} = 0, S_{\p{B}} = 0} + \ba{\p A\p B | S_{\p{A}} = 0, S_{\p{B}} = 1} + \ba{\p A\p B | S_{\p{A}} = 1, S_{\p{B}} = 0} - \ba{\p A\p B | S_{\p{A}} = 1, S_{\p{B}} = 1} \leq 2 \eq \label{eq:CHSH_ineq_orig} \]
    Non-trivial Bell inequalities such as the CHSH inequality are capable of distinguishing the non-classical nature of distributions beyond the constraints implied by observable d-separation relations; the inequalities presented in \cref{sec:found_inequalities} are also of this type.

>>>>

\begin{document}
    \title{Causal Compatibility Inequalities Admitting of Quantum Violations in the Triangle Structure}
    \author{Thomas C. Fraser}
    \email{tcfraser@tcfraser.com}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada, N2L 2Y5}
    \affiliation{University of Waterloo, Waterloo, Ontario, Canada, N2L 3G1}
    \author{Elie Wolfe}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada, N2L 2Y5}
    % \email{ewolfe@perimeterinstitute.ca}
    \date{\today}
    \begin{abstract}
        It has long been recognized that certain quantum correlations are incompatible with particular assumption about classical causal structure. Given a causal structure of unknown classicality, the presence of such correlations certifies the non-classical nature of the causal structure in a device independent fashion. In scenarios where all parties share a common resource, these non-classical correlations are also known as non-local correlations. Any constraint satisfied by all correlations which \textit{are} classically compatible with a given causal structure defines a causal compatibility criterion. Such criteria were recently derived for the Triangle structure [\href[pdfnewwindow]{http://arxiv.org/abs/1609.00672}{arXiv:1609.00672}] in the form of polynomial inequalities, begging the question: do any of those inequalities admit violation by quantum correlations? Numerical investigation suggests not, and we further conjecture that the set of correlations admitted by the classical Triangle structure is equivalent to the set of correlations admitted by its quantum generalization whenever the three observable variables are binary. Our main contribution in this work, however, is the derivation of new causal compatibility inequalities for the Triangle structure which \textit{do} admit quantum violation. This provides the first robust-to-noise witness of quantum correlations in the Triangle structure. We conclude by considering the possibility of quantum resources potentially qualitatively different from those known previously.
    \end{abstract}
    \maketitle
    % \tableofcontents
    % \setlength{\parskip}{1em}

@@

    From a foundational prospective, the most robust demonstrations of quantum phenomena with no classical emulation have involved Bell inequalities~\cite{Bell_1964,Brunner_2013}. Originally, Bell inequalities were derived as a way to show that no hidden variable theory could ever account for quantum mechanics; in this sense Bell inequalities are a response to the famous EPR paradox~\cite{EPR_Orig}. The enumeration of Bell inequalities has since become a widespread systematic way of demonstration the \textit{non-classicality} of a given observation. More recently it has been appreciated that Bell inequalities can understood as consequences of \term{causal inference}~\cite{Wood_2012}. Causal inference is concerned with classifying observations into those which can and cannot be explained by a hypothesized causal structure. The abstract nature of causal inference is responsible for its presence in numerous scientific fields including machine learning and biology~\cite{Pearl_2009,Pearl_2009_tr}. Causal compatibility inequalities, such as Bell and Instrumental inequalities \cite{pearl1995instrumental,bonet2001instrumental,evans2012graphical}, characterize the space of observations that are compatible with a hypothesized causal structure, albeit the characterization offered by practically derivable inequalities is often only an approximation. To derive the traditional Bell inequalities from causal inference one starts with a (classical) causal structure known as the Bell structure, as depicted in \cref{fig:bell_structure}. The fundamental Bell structure involves non-communicating parties making measurements on some hidden shared resource $\la$, where the measurement outcomes ($A, B$) are presumed to be stochastic functions of the the local choices of measurement settings ($S_A, S_B$) and the shared resource $\la$. Quantum non-classicality in the Bell structure has been thoroughly studied since Bell's original work~\cite{Brunner_2013}. More complex structures, however, such as the correlation scenarios proposed by Fritz~\cite{Fritz_2012,Fritz_2014}, are much less understood. Here we investigate one particular causal structure: the Triangle structure (\cref{fig:triangle_structure}).

    The Triangle structure (\cref{fig:triangle_structure}) is a causal structure comprised of $3$ parties labeled $A, B, C$ arranged in a triangular configuration while pair-wise sharing hidden/latent variables $X, Y, Z$. It has been extensively studied previously (see~\cite[Fig. 1]{Steudel_2010},~\cite[Fig. 6]{Chaves_2014},~\cite[Fig. 8]{Branciard_2012},~\cite[Fig. 8, App. E]{Henson_2014},~\cite[Fig. 3]{Fritz_2012},~\cite[Fig. 4]{Weilenmann_2016},~\cite[Fig. 1]{Inflation}, $\ldots$). An overview of some milestone results is provided in \cref{sec:triangle_structure}. Identifying causal compatibility inequalities for this configuration has been seen as particularly challenging~\cite{Branciard_2012}. Further identifying causal compatibility inequalities of such high resolution that such inequalities can be violated by quantum-accessible distributions has remained out-of-reach for the Triangle structure.

    This work is the first, therefore, to find causal compatibility inequalities for the Triangle structure that are able to be violated by quantum-accessible distributions. This accomplishment was made possible through the combination of two previous developments. First, the insight of \citet{Fritz_2012} regarding the ability to re-interpret the Bell structure as a portion of the Triangle structure. Second, the new framework for solving causal inference problems developed by Wolfe \emph{et al.}~\cite{Inflation} called \term{The Inflation Technique}. Ultimately, this work serves as a validation that the Inflation Technique is sensitive enough to offer new insights into quantum non-classicality \cite{Navascues_2017}. Moreover, these new inequalities offer an avenue for recognizing previously unknown forms of non-classicality. The authors attempts to find such novel resources were met with only partial success, suggesting clear refinements for future exploration.

    The manuscript is organized into the following sections: \Cref{sec:causal_compatibility} recalls important notions from causal inference theory and sets up the notation to be used. \Cref{sec:bell_structure} offers a summary of the popular Bell structure and associated inequalities. \Cref{sec:triangle_structure} discusses the Triangle structure and provides an overview of existing research; identifying its stark differences from the Bell structure and motivating why the Triangle structure is worth studying. \Cref{sec:fritz_distribution} defines and discusses a singularly-quantum correlation first conceived of by \citet{Fritz_2012}, which we term the \term{Fritz distribution}. The Fritz distribution first was proven to be non-classical without the use of inequalities. Here we offer a direct proof using inequalities, with the advantage that our proof is robust to noise. A sample of such inequalities are presented in \cref{sec:found_inequalities}, specifically \cref{eq:ww_ineq,eq:web_inequality,eq:symmetric_web_inequality}. In \cref{sec:violations_noise} we quantify the robustness to noise of the Fritz distribution, and we numerically optimize our derived inequalities over quantum-accessible distributions (using qubits) in an effort to seek out new new flavors of of non-classicality in the Triangle structure. Additionally, the Fritz distribution is subjected to noise in order to measure the robustness of the derived inequalities. \Cref{sec:conclusions} concludes with proposed avenues for further research.

    \Cref{sec:main_appendix} briefly summaries the Inflation Technique in the specific context of this work. Although the summary presented here is designed to be self-standing, a much more pedagogical introduction is offered by the original work~\cite{Inflation}. \Cref{sec:deriving_inequalities} demonstrates how the Inflation Technique was used to derive the causal compatibility inequalities for the Triangle structure which admit violation by quantum-accessible distributions.

@@

    Henceforth, we will utilize a number of familial notions from graph theory and denote them accordingly. The \term{parents of a node} $n \in \nodes$ are all nodes which point directly into $n$, i.e. $\Pa[\graph]{n} \defined \bc{m \mid m \to n}$. Similarly defined are the \term{children of a node} $\Ch[\graph]{n} \defined \bc{m \mid n \to m}$. Recursively defined are the \term{ancestors of a node} $\An[\graph]{n} \defined \bigcup_{i\in\mathbb{N}} \Pa[\graph][i]{n}$ where $\Pa[\graph][i]{n} \defined \Pa[\graph]{\Pa[\graph][i-1]{n}}$ and $\Pa[\graph][0]{n} = n$ and the \term{descendants of a node} $\De[\graph]{n} \defined \bigcup_{i\in\mathbb{N}} \Ch[\graph][i]{n}$ where $\Ch[\graph][i]{n} \defined \Ch[\graph]{\Ch[\graph][i-1]{n}}$ and $\Ch[\graph][0]{n} = n$. Finally, we extend this notation to a subset of nodes $N \subseteq \nodes$ by performing a union over elements. As an example, the parents of the nodes $N \subseteq \nodes$ are denoted $\Pa[\graph]{N} = \bigcup_{n \in N} \Pa[\graph]{n}$.

@@

    Let us now formalize the notion of compatibility between a causal structure $\graph$ and a probability distribution $\prob[\nodes]$ defined over the nodes of $\graph$. Intuitively, the absence of certain causal connections within $\graph$ should limit the set of possible observations that can be made on $\graph$. Specifically, the causal structure $\graph$ hypothesizes that if each variable $n \in \nodes$ is causally influenced by its parents $\Pa[\graph]{n}$. Therefore, if each variable is conditioned on its parentage, the probability distribution $\prob[\nodes]$ should factorize accordingly:
    \[ \prob[\nodes] = \prod_{n \in \nodes} \prob[n \mid \Pa[\graph]{n}] = \prob[n_1 \mid \Pa[\graph]{n_1}] \times \cdots \times \prob[n_k \mid \Pa[\graph]{n_k}] \eq \label{eq:compatibility_as_factorization}\]
    % , except its own descendants when conditioned over $\Pa[\graph]{n}$. This is known as the \term{local Markov property}:\footnote{Here $A \indep B \mid C$ denotes statistical independence between $A$ and $B$ conditional on $C$ where $A,B,C \subseteq \nodes$, i.e. for a given probability distribution $\prob[\nodes]$, $\prob[AB \mid C] = \prob[A \mid C]\prob[B \mid C]$.}
    % \[ n \indep \br{\nodes \setminus \De[\graph]{n}} \mid \Pa[\graph]{n} \quad \forall n \in \nodes \eq \label{eq:local_markov_property} \]
    If a given distribution $\prob[\nodes]$ defined over \textit{all} of the nodes $\nodes$ of $\graph$ satisfies \cref{eq:compatibility_as_factorization}, then $\prob[\nodes]$ is said to be \term{compatible with $\graph$}. The conditional distributions in \cref{eq:compatibility_as_factorization} (i.e. $\{\prob[n \mid \Pa[\graph]{n}] \mid n \in \nodes\}$) are referred to as a set of \term{causal parameters for $\graph$}. If a distribution $\prob[\nodes]$ cannot be factorized according to \cref{eq:compatibility_as_factorization}, $\prob[\nodes]$ is said to be \term{incompatible} with $\graph$. 
    % The compatibility of a distribution $\prob[\nodes]$ with a causal structure $\graph$ pursuant to \cref{eq:compatibility_as_factorization} is equivalent to the requirement that $\prob[\nodes]$ obeys all of the conditional independence constraints implied by the \textit{d-separation relations} between variables/nodes in $\graph$~\cite{Pearl_2009}.
     When one is specified with a joint distribution $\prob[\nodes]$ defined over all nodes of a causal structure $\graph$, it is possible to completely determine whether or not $\prob[\nodes]$ is compatible with $\graph$ by computing the causal parameters induced by $\prob[\nodes]$ and checking the equality of \cref{eq:compatibility_as_factorization}.

    A challenge, however, is presented when one is supplied with a partial observation, i.e. a joint distribution $\prob[\nodes_{O}]$ where $\nodes_{O} \subset \nodes$ is some subset of variables referred to as the \term{observable nodes} $\nodes_O$. In such cases, $\prob[\nodes_{O}]$ does not induce a unique set of causal parameters for $\graph$ and \cref{eq:compatibility_as_factorization} can not be verified by direction calculation. Instead, compatibility between $\graph$ and $\prob[\nodes_{O}]$ depends on the existence or non-existence of a set of causal parameters for $\graph$ such that $\prob[\nodes_{O}] = \prod_{n \in \nodes_{O}} \prob[n \mid \Pa[\graph]{n}]$. The complementary, unobservable nodes are termed \term{latent nodes} $\nodes_L = \nodes \setminus \nodes_{O}$, and should be understood as hidden random variables that are unknowable either by some fundamental process or cannot be measured due to other limitations.

    There are several approaches to tackling the compatibility problem when dealing with latent variables; there are two common approaches worth mentioning here. The first is to recognize that many equality constraints are implied by the causal structure, including conditional independence relations and so-called Verma constraints among others; see Refs. \cite{evans2015margins,evans2017markov} for thorough treatments. The failure to satisfy an equality constraint immediately disqualifies $\prob[\nodes_{O}]$ from being compatible with $\graph$. Equality constraints are easily derived given a causal structure, and checking equality-constraints satisfaction is the minimalistic algorithm which powers the overwhelming majority of practical causal inference hypothesis testing in the fields of machine learning and artificial intelligence. In quantum theory, however, we require strong, more sensitive, causal inference techniques. This is because the equality constraints satisfied by classical correlations (compatible with some classical causal structure) are also all satisfied by quantum correlations (generated by the quantumified causal structure in which latent variables are replaced with shared entanglement) \cite{Henson_2014}. 
    Our focus, therefore, in on deriving \emph{inequality} constraints (over $\prob[\nodes_{O}]$ and its marginals) implied by a causal structure, which we term \term{causal compatibility inequalities}\footnote{We refer to these inequalities as ``causal compatibility inequalities'' instead of Bell inequalities for two reasons. Firstly, ``Bell inequalities'' usually are associated specifically with the Bell structure. Secondly, the inequalities derived in this work are fundamentally distinct from a typical Bell inequality in that these inequalities are \textit{polynomial} over $\prob[\nodes_{O}]$ instead of \textit{linear}.}. For some causal structures, the equality constraints associated with it are sufficient to perfectly characterize the distributions genuinely compatible with it; for others, however, inequality constraints are also important. Causal structures for which inequality constraints are relevant have been termed \textit{interesting}~\cite{Henson_2014}, and such structures include the Instrumental structure, the Bell scenario, and the Triangle scenario studied here, among infinitely many others. 
    Herein we use the Inflation Technique~\cite{Inflation} to find causal compatibility inequalities; \Cref{sec:main_appendix} discusses the Inflation Technique as applied in this work.

     probabilistic inequalities defined over $\prob[\nodes_{O}]$ and its marginals. We term these constraints \term{causal compatibility inequalities}\footnote{We refer to these inequalities as ``causal compatibility inequalities'' instead of Bell inequalities for two reasons. Firstly, ``Bell inequalities'' usually are associated specifically with the Bell structure. Secondly, the inequalities derived in this work are fundamentally distinct from a typical Bell inequality in that these inequalities are \textit{polynomial} over $\prob[\nodes_{O}]$ instead of \textit{linear}.}. \Cref{sec:main_appendix} provides a brief discussion of existing techniques used to derive these inequalities.

    If a probability distribution $\prob[\nodes_{O}]$ happens to \textit{violate} any causal compatibility inequality, then that distribution deemed \textit{incompatible}.  Conversely, however, a singular inequality can only be used to prove that a given distribution is \textit{incompatible}; a single inequality cannot certify compatibility. A \textit{complete characterization} of compatibility consists of a complete set of all valid causal compatibility inequalities such that satisfaction of the entire set certifiescompatibility. Currently however, it is unknown how to obtain a complete characterization for any given causal structure, including the Triangle structure.

    From the perspective of identifying quantum non-classicality, a causal structure $\graph$ adopts the role of a classical hypothesis. Therefore, non-classicality becomes synonymous with incompatibility: if a distribution $\prob[\nodes_{O}]$ is incompatible, then it is non-classical. Henceforth, we will use these two terms interchangeably. From a resource standpoint, if the non-classicality of $\prob[\nodes_{O}]$ with $\graph$ can be witnessed by an inequality $I$, but nevertheless $\prob[\nodes_{O}]$ can be implemented using quantum states and measurements while otherwise respecting the causal relations of $\graph$, then the causal compatibility inequality $I$ represents a task or game where quantum resources out-perform classical resources relative to $\graph$.

    \section{Bell Structure}
    \label{sec:bell_structure}
    This section aims to define the Bell causal structure and to review some of the traditional witnesses used to asses the classicality (or lack thereof) of distributions relative to it. The purpose of this section is to equip readers with the pertinent background and also to draw comparisons between the advancements made toward understanding non-classicality of the Bell structure versus analogous results obtained for Triangle structure later in this work.

    The bipartite Bell structure (\cref{fig:bell_structure}) refers to an iconic causal structure involving two distant parties who observe the outcomes of local measurements as random variables $A (B)$ determined by their individual measurement settings $S_A (S_B)$, and where the parties are also presumed to be commonly informed by some shared latent resource $\la$~\cite{Brunner_2013}. The resulting observable correlations hence naturally form a conditional probability distribution $\prob[AB|S_AS_B]$. Subject to the notions of compatibility presented in \cref{sec:causal_compatibility}, this distribution is compatible with \cref{fig:bell_structure} if and only if there exists some functional specification of $\la$ and causal parameters $\prob[A|S_A,\la]$, $\prob[B|S_B,\la]$ such that 
%    there exists a distribution $\prob[AB|S_AS_B,\la]$ such that $\prob[AB|S_AS_B] = \sum_{\la} \prob[AB|S_AS_B,\la]$ and also satisfies the following implementation of \cref{eq:compatibility_as_factorization}:
\begin{align}\label{eq:bellcompat}
 \prob \text{ is classically compatible with \cref{fig:bell_structure}} \quad\text{iff}\quad \prob[AB|S_AS_B] = \sum_{\la} \prob[A|S_A,\la]\prob[B|S_B,\la]\,.
\end{align}
    Any observed distribution $\prob[AB|S_AS_B]$ which fails to be explained by the classical causal hypothesis of the Bell structure as defined by \cref{eq:bellcompat} is appropriately termed non-classical. Often, distributions incompatible with the Bell structure are referred to as non-local because the Bell structure markedly lacks causal influence from one party to another. In particular,~\citet{Bell_1964} demonstrated that there exists distributions that are non-classical yet are attainable via local measurements on a shared quantum resource. 
    %If one does not wish to accept fine-tuning arguments, one must accept that quantum correlations can be non-local as no variation to the latent nodes can explain the same correlations~\cite{Wood_2012}.\footnote{This is unlike the Triangle structure wherein a global yet latent common cause is always an acceptable explanation for any correlation and does require fine-tuning. Therefore, we never refer to non-classical correlations in Triangle Structure as non-local.}
